### 2023-03-03 manuscript.Rmd

Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable data sets, and research software.
We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
Why are research software contributions important scientific contributions?
We would like to respond with a quote from the Science Code Manifesto [@manifesto]:
"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
and
"Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice [@baxter2012research].
In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
rigor means implementing high standards and best practices for ensuring transparency, correctness, and re-usability of a piece of software.
By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).

@@

Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
We appreciate the importance of assessing maintenance as part of rigor.
To assess this, we propose to provide a simple check box, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.

@@

# rigor for software implementations

>>>>

Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable data sets, and research software.
We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
Why are research software contributions important scientific contributions?
We would like to respond with a quote from the Science Code Manifesto [@manifesto]:
"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
and
"Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice [@baxter2012research].
In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
Rigor means implementing high standards and best practices for ensuring transparency, correctness, and re-usability of a piece of software.
By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).

@@

Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
To assess active maintenance as an aspect of rigor, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.

@@

# Assessing Rigor
